---
title: "The main ideas behind the MocapMineR package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The main ideas behind the MocapMineR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

`MocapMineR` is an R package dedicated to the analysis of motion capture data. It was initially developed by Lou-Anne Guillotel and myself (Sébastien Lê) to analyze movements from blind Tae Kwon Do practitioners. The main idea, was to find a way to correct these practitioners automatically, based on graphical indicators. Actually, this package works for any kind of motion capture data. To use this package you will need two data sets. A first one, in which you store the coordinates of joints as a function of time, a second one in which you describe the structure of interest, *i.e.* the dependencies between the joints that define the segments of the structure.

# A first step: vizualization of the structure as a function of time with the `skeleton` function

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(MocapMineR)
```

The *human* data set depicts the dependencies between the segments that compose the structure. It is constituted of three columns, a first one (*segment*) with the names of the segments, the two others with the names of the joints at the extremities of the segments (*extr1* and *extr2*).

```{r, chunk00, echo=-1}
options(width = 100)
data(human)
head(human)
```

Then, the *gaetan_apchagi* data set contains the coordinates of the joints as a function of time. The first column is the moment when the information have been recorded. The second column is the joint that was measured. Columns 3 to 5 correspond to the dimensions that have been measured. Column 6 is the index of the frame. Column 7 is the number of frames per second and the last column is the name of the practitioner.

```{r, chunk0, echo=-1}
options(width = 100)
data(gaetan_apchagi)
head(gaetan_apchagi)
```

The *skeleton* function allows you to vizualize the movement you have recorded. The parameters *num.joint*, ..., *num.name* depend on the *gaetan_apchagi* data set.

```{r, chunk1, echo=-1}
options(width = 100)
data(gaetan_apchagi)
data(human)
skeleton(joint=gaetan_apchagi, structure=human, num.joint=2, 
         num.frame=6, num.x=3, num.y=4, num.name=8)
```

You may also want to highlight a particular segment or body part.

```{r, chunk11, echo=-1}
options(width = 100)
data(gaetan_apchagi)
data(human)
skeleton(joint=gaetan_apchagi, structure=human, num.joint=2, 
         num.frame=6, num.x=3, num.y=4, num.name=8,
         body.part = "tibia_r",color.part = "orange")
```

You can also want to represent a particular frame, and highlight a particular segment or body part.

```{r, chunk2, echo=-1}
options(width = 100)
data(gaetan_apchagi)
data(human)
skeleton(joint=gaetan_apchagi, structure= human, num.joint=2,num.frame=6,
         num.x=3, num.y=4, num.name=8, body.part = "tibia_r",frame.index=25,
         color.part = "orange")
```




